---
title: "Exercise 2 Dealing With Missing Values"
author: "Winnie Han"
date: "24 July 2017"
output: html_document
---



```{r echo=TRUE, message=FALSE}
library(dplyr)
library(tidyr)
```

### 0: Load the data in RStudio
```{r echo=TRUE, message=FALSE}
titanic_original <- read.csv("titanic_original.csv")
```

### 1: Port of embarkation
```{r echo=TRUE, message=FALSE}
titanic_original$embarked[titanic_original$embarked == ""] <- "S" 
```

### 2: Age
```{r echo=TRUE, message=FALSE}
titanic_original$age[is.na(titanic_original$age)] <- round(mean(titanic_original$age, na.rm = TRUE))
```
> Task 2.2: Think about other ways you could have populated the missing values in the age column. Why would you pick any of those over the mean (or not)?
>
  + *Mean +/- 0.5* - as other estimated ages are in the form of **xx.5**.
  + *Median*, OR *Mean by Sex (male vs female)*.
  + All results are very close *(between 28-31)*, so without any further analysis, using the **Mean** seems to be the best way forward for now.


### 3: Lifeboat
```{r echo=TRUE, message=FALSE}
titanic_original$boat[titanic_original$boat == ""] <- NA
```

### 4: Cabin
```{r echo=TRUE, message=FALSE}
titanic_clean <- titanic_original %>% 
  mutate(has_cabin_number = ifelse(cabin == "",0,1))
```
> Does it make sense to fill missing cabin numbers with a value?
>
  + Unless we can identify how the cabin numbers are coded, it doesn't make any sense to guess and fill in the missing values.
>
> What does a missing value here mean?
>
  + Either the cabin number is *unknown* OR the passenger didn't have a dedicated cabin - in which case, just the **has_cabin_number** flag should still help with further analysis.


### 6: Submit the project on Github
```{r echo=TRUE, message=FALSE}
write.csv(titanic_clean, file = "titanic_clean.csv", row.names = FALSE)
```
