---
title: "Run Analysis Read Me"
author: "Winnie Han"
date: "1 August 2017"
output: html_document
---

#### Load required packages 
```{r echo=TRUE, message=FALSE}
library(dplyr)
library(tidyr)
```

### 0: Load the data in RStudio 

#### Variable names for *test* and *train* data sets, rename columns, make rows unique.
```{r echo=TRUE, message=FALSE}
features <- read.table("UCI HAR Dataset/features.txt") 
features <- features %>% 
  setNames(c("Key", "Name")) %>% 
  mutate(Name = make.names(Name, unique = TRUE)) 
```

#### *Activity labels* and *activity names*, rename columns.
```{r echo=TRUE, message=FALSE}
act_labels <- read.table("UCI HAR Dataset/activity_labels.txt") 
act_labels <- act_labels %>% 
  setNames(c("ActivityLabel", "ActivityName"))
```

#### *Activity labels* for *test* data, rename column.
```{r echo=TRUE, message=FALSE}
test_act <- read.table("UCI HAR Dataset/test/y_test.txt") 
test_act <- test_act %>% 
  setNames(c("ActivityLabel")) 
```

#### *Subjects* for *test* data, rename column.
```{r echo=TRUE, message=FALSE}
test_subject <- read.table("UCI HAR Dataset/test/subject_test.txt") 
test_subject <- test_subject %>%
  setNames(c("Subject"))
```

#### *Test* data, rename columns using *features*, add *activity labels* and *subjects* as new variables.
```{r echo=TRUE, message=FALSE}
test <- read.table("UCI HAR Dataset/test/X_test.txt") 
test <- test %>% 
  setNames(features$Name) %>% 
  mutate(ActivityLabel = test_act$ActivityLabel, Subject = test_subject$Subject)
```

#### *Activity labels* for *train* data, rename column.
```{r echo=TRUE, message=FALSE}
train_act <- read.table("UCI HAR Dataset/train/y_train.txt") 
train_act <- train_act %>% 
  setNames(c("ActivityLabel")) 
```

#### *Subjects* for *train* data, rename column.
```{r echo=TRUE, message=FALSE}
train_subject <- read.table("UCI HAR Dataset/train/subject_train.txt") 
train_subject <- train_subject %>%
  setNames(c("Subject"))
```

#### *Train* data, rename columns using *features*, add *activity label* and *subject* as new variables.
```{r echo=TRUE, message=FALSE}
train <- read.table("UCI HAR Dataset/train/X_train.txt")
train <- train %>% 
  setNames(features$Name) %>% 
  mutate(ActivityLabel = train_act$ActivityLabel, Subject = train_subject$Subject)
```

### 1: Merge data sets 
#### Merge *test* and *train* datasets into a new dataset.
```{r echo=TRUE, message=FALSE}
ucihar <- bind_rows(test, train)
```

### 2: Mean and standard deviation
#### Add *activity mean* and *activity standard deviation* as new variables for each observation.
```{r echo=TRUE, message=FALSE}
ucihar <- ucihar %>% 
  mutate(ActivityMean = rowMeans(ucihar), ActivityStdDev = apply(ucihar, 1, sd)) 
```

### 3: Add new variables 
#### *Activity label* already added to *test* and *train* separately prior to merge, add *activity name*.
```{r echo=TRUE, message=FALSE}
ucihar <- left_join(ucihar, act_labels, by = "ActivityLabel")
```

### 4: Create tidy data set 
#### Group by *activity label / name* and *subject*, calculate the mean for all other variables.
#### Assign the result to a new data set.
```{r echo=TRUE, message=FALSE}
run_analysis <- ucihar %>% 
  group_by(ActivityLabel, ActivityName, Subject) %>% 
  summarise_each(funs(mean(., na.rm = TRUE))) 
```

### 5: Submit on github 
#### Write the tidy data set created above to a new CSV file.
```{r echo=TRUE, message=FALSE}
write.csv(run_analysis, file = "run_analysis.csv", row.names = FALSE)
```